Convergence rate of MAP estimates \\ for the exponential family

### 2021-10-14 aistats.tex
Move some related work to intro

%\textcolor{red}{This comes a bit from nowhere. Maybe phrase it like (its not sexy wirting here, just the idea of the structure) "The measure of closeness between theta and its estimate hat theta is an active area of research. We distinguish two way to do that: 1) distance in the parameter space and 2) distance between their associated distribution. In the first case, the analysis use norms blabla. In the second, we use KL"}
{\bf Metrics.}
Once we have an estimator $\hat \nat$, we wish to assess its quality:
how close is it from some optimal $\nat^*$.
This is necessary, for instance to choose between 2 estimators (model selection).
We distinguish two ways of doing that.
{\bf 1)} distance in parameter space $d(\nat,\nat^*)$. This is the focus of \emph{parameter estimation}, yielding results such as the asymptotic efficiency of the MLE via the Cramer-Rao lower-bound \citep{aitken1942estimation}.
{\bf 2)} distance between distributions, as studied in \emph{density estimation}.
For this purpose, the Kullback-Leibler (KL) divergence $\KL\paren{p_{\nat^*} || p_{\nat} }$  arises naturally from information theory,
but its lack of robustness to misspecification\footnote{For $p$ and $q$ continuous densities of the ground truth and a model,
$\KL(p||q) = +\infty$ if $q(x)=0$ for some $x$ such that $p(x)>0$.
}
has led statisticians to study other distances, such as the $L^2$ norm \citep[e.g. \S1.2]{tsybakov2009introduction}.
As we will see, with exponential families, the KL divergence is also a Bregman divergence between parameters, thus drawing a connection between these two lines of research.
Hence the fundamental question:
%\textcolor{red}{The sample complexity of the quality of the estimator is well known when we use norms between parameters, in opposition to the case where we measure the distance between distribution. }
% RLP : that is not so true. the citation tsybakov does gie bound for non-parametric estimation with kernels in L^2 distance.
%\begin{center}
%	\framebox{
%		\parbox[c][]{0.9\linewidth}{
%			\begin{center}
%			\emph{Can we upper bound  the expected value of the KL divergence for MLE or MAP?}
%		\end{center}
%		}
%	}
%\end{center}

\begin{equation}
\boxed{
\begin{aligned}
	\textit{Can we upper}&\textit{ bound  the expected value of} \\
	&\KL(p_{\nat^*} || p_{\hat \nat_n^\text{MLE/MAP} }) \ ?
\end{aligned}
}
\tag{$\star$}
\label{problem}
\end{equation}
% RLP : that is a dirty hack, but I would like to be able to reference the problem somehow.

@@

\paragraph{Parameter Estimation.}
Statistics has a plethora of asymptotic results \citep{vdv1998asymptotic}.
In particular for sum of independent variables such as~\eqref{eq:defMLE}, large deviations theory \citep{varadhan1984large} characterizes the decay of the tail.

\paragraph{Density Estimation.}
Since the KL explode to $\infty$ if the model is unable to give weight to some samples, a lot of this work has been devoted to distances, such as $L^2$ \citep{tsybakov2009introduction}, $L^1$ \citep{devroye2001combinatorial} or more recently Hellinger distance \citep{baraud2017new}.
Contrary to the KL, these distance are symmetric and behave smoothly.

>>>>

%\textcolor{red}{This comes a bit from nowhere. Maybe phrase it like (its not sexy wirting here, just the idea of the structure) "The measure of closeness between theta and its estimate hat theta is an active area of research. We distinguish two way to do that: 1) distance in the parameter space and 2) distance between their associated distribution. In the first case, the analysis use norms blabla. In the second, we use KL"}
{\bf Metrics.}
Once we have an estimator $\hat \nat$, we wish to assess its quality:
how close is it from some optimal $\nat^*$.
This is necessary, for instance to choose between 2 estimators (model selection).
We distinguish two ways of doing that.
{\bf 1)} distance in parameter space $d(\nat,\nat^*)$. This is the focus of \emph{parameter estimation}, yielding results such as the asymptotic efficiency of the MLE via the Cramer-Rao lower-bound \citep{aitken1942estimation} and a wealth of asymptotic results \citep{vdv1998asymptotic}.
In particular for sum of independent variables such as~\eqref{eq:defMLE}, large deviations theory \citep{varadhan1984large} characterizes concentration phenomena.
{\bf 2)} distance between distributions, as studied in \emph{density estimation}.
For this purpose, the Kullback-Leibler (KL) divergence $\KL\paren{p_{\nat^*} || p_{\nat} }$  arises naturally from information theory,
but its lack of robustness to misspecification\footnote{
For $p$ and $q$ continuous densities,
$\KL(p||q) = +\infty$ if $\exists x, q(x)=0 \&p(x)>0$. %Since exponential families have a positive density over their whole input set, this is not a concern for us, except if we misspecified the input set.
}
has led statisticians to study symmetric, better-behaved distances, such as the $L^2$ norm \citep[e.g. \S1.2]{tsybakov2009introduction} , the $L^1$ norm \citep{devroye2001combinatorial} or more recently the Hellinger distance \citep{baraud2017new}.
With exponential families, the KL divergence is also a Bregman divergence between parameters (see \S\ref{sec:problem}), thus drawing a connection between these two lines of research.
Hence the fundamental question:
%\textcolor{red}{The sample complexity of the quality of the estimator is well known when we use norms between parameters, in opposition to the case where we measure the distance between distribution. }
% RLP : that is not so true. the citation tsybakov does gie bound for non-parametric estimation with kernels in L^2 distance.
%\begin{center}
%	\framebox{
%		\parbox[c][]{0.9\linewidth}{
%			\begin{center}
%			\emph{Can we upper bound  the expected value of the KL divergence for MLE or MAP?}
%		\end{center}
%		}
%	}
%\end{center}

\begin{center}
\begin{equation}
\boxed{
\begin{aligned}
	\textit{Can we upper}&\textit{ bound  the expected value of } \\
	&\KL(p_{\nat^*} || p_{\hat \nat_n^\text{MLE/MAP} }) \ ?
\end{aligned}
}
\tag{$\star$}
\label{problem}
\end{equation}
\end{center}
% RLP : that is a dirty hack, but I would like to be able to reference the problem somehow.

@@
